Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-04-27 doc/manuscript/ESSDmanuscript.tex
knit with updated table

\textbf{Abstract.} Forests are critical for climate change mitigation
and consitute a substantial portion of planned emissions reductions
under the 2015 Paris Agreement. Yet, the efficacy of greenhouse gas
mitigation planning and reporting is dependent upon the quality of
available emission factors data, including forest carbon (C) stocks and
changes therein. Tens of thousands of relevant forest C estimates have
been published, yet are not readily accesible to the practitioners
compiling national greenhouse gas inventories. Many of these data have,
however, been compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
accounting if made available through the Emission Factor Database (EFDB)
of the International Panel on Climate Change (IPCC). Here, we develop
and document a process for semi-automated transfer of data from ForC
into the EFDB, assess the data available and transferred to date, and
provide recommendations for improving forest data collection, analysis,
and reporting to improve accounting of forest-sector greenhoouse gas
emissions and removals. We begin by reconciling terminology and mapping
ForC fields into EFDB. This process required some updates to the ForC
database structure, leading to the release of a new version of ForC
(v4.0; described here). At the time of writing, ForC contained
\textasciitilde17190 independent records that would be relevant to EFDB,
1214 of which have been submitted to date. Among the data in ForC, there
is disproportionate representation of biomass (particularly aboveground)
stocks, with far fewer records for dead organic matter and soil carbon,
and relatively few or no records for net annual increments or C fluxes
into (gains) or out of (losses) the IPCC-defined C pools. Geographic
representation is also quite uneven, with the highest densities of
relevant records in temperate forests, and with relatively scant
representation of tropical forests in Africa and Asia. ForC represents a
diversity of stand ages, although records for young stands are primarily
limited to C stocks, as opposed to net increments of fluxes. This
distribution of records is generally reflected in the subset of records
that have been submitted to EFDB to date. In the future, forest C
estimates in EFDB can be improved through targeted research to fill
critical gaps, reporting of information required by IPCC, and continued
submission of data from scientific publications to the EFDB.

@@

As of April 27, 2022, ForC (v4.0) contained 32550 independent records
(39856 total), 17190 of which were for variables relevant to EFDB (Fig.
1). These records were distributed across all forested continents and
ecozones, with particularly high concentrations in
\emph{{[}ecozone(s){]}} and low concentrations in
\emph{{[}ecozone(s){]}} (Fig. 2). The most widely represented forest
type was \emph{{[}type{]}}, followed by \emph{{[}type{]}} and
\emph{{[}type{]}} (Fig. 3).

@@

\begin{longtable}[t]{l|l|l|l|l}
\caption{\label{tab:table_variables}\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.}}\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\endfirsthead
\caption[]{\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} \textit{(continued)}}\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\endhead
\textbf{Biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass & 1095 & 847 & 93 & 48\\
\hline
delta.biomass & 0 & 0 & 0 & 0\\
\hline
NPP\_woody & 136 & 93 & 0 & 0\\
\hline
woody.mortality & 0 & 0 & 0 & 0\\
\hline
\textbf{Aboveground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_ag & 9038 & 7724 & 1232 & 693\\
\hline
biomass\_ag\_woody & 460 & 366 & 10 & 10\\
\hline
biomass\_ag\_foliage & 601 & 502 & 49 & 27\\
\hline
delta.agb & 166 & 128 & 123 & 123\\
\hline
ANPP\_woody & 299 & 242 & 0 & 0\\
\hline
ANPP\_foliage & 0 & 0 & 0 & 0\\
\hline
woody.mortality\_ag & 113 & 61 & 17 & 17\\
\hline
\textbf{Belowground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_root & 4629 & 4180 & 123 & 55\\
\hline
biomass\_root\_fine & 931 & 594 & 18 & 18\\
\hline
biomass\_root\_coarse & 599 & 410 & 12 & 7\\
\hline
delta.biomass\_root & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_coarse & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_fine & 0 & 0 & 0 & 0\\
\hline
woody.mortality\_root & 0 & 0 & 0 & 0\\
\hline
BNPP\_root\_fine & 489 & 333 & 0 & 0\\
\hline
BNPP\_root.turnover\_fine & 91 & 56 & 0 & 0\\
\hline
BNPP\_root\_coarse & 329 & 250 & 0 & 0\\
\hline
\textbf{Dead wood} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
deadwood & 438 & 303 & 94 & 61\\
\hline
deadwood\_standing & 152 & 120 & 17 & 17\\
\hline
deadwood\_down & 424 & 368 & 51 & 27\\
\hline
delta.deadwood & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_standing & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_down & 0 & 0 & 0 & 0\\
\hline
R\_het\_deadwood & 0 & 0 & 0 & 0\\
\hline
\textbf{Litter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
O.horizon & 38 & 38 & 38 & 38\\
\hline
delta.O.horizon & 4 & 4 & 4 & 4\\
\hline
litter & 30 & 30 & 23 & 23\\
\hline
delta.litter & 0 & 0 & 0 & 0\\
\hline
NPP\_litter & 0 & 0 & 0 & 0\\
\hline
R\_het\_litter & 167 & 143 & 0 & 0\\
\hline
\textbf{Soil organic matter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
SOM / SOC & 693 & 398 & 81 & 46\\
\hline
delta.SOM / delta.SOC & 0 & 0 & 0 & 0\\
\hline
R\_het\_soil & 0 & 0 & 0 & 0\\
\hline
\textbf{TOTAL} & \textbf{20922} & \textbf{17190} & \textbf{1985} & \textbf{1214}\\
\hline
\end{longtable}
\endgroup{}

@@

As of April 27, 2022, we had reviewed or added 1985 EFDB-relevant
records, 1214 records of which were sent to EFDB, and 73 of which have
been reviewed, accepted, and posted (Figs. 2-3, Table 2).
\emph{{[}DETAILS{]}}

>>>>

\textbf{Abstract.} Forests are critical for climate change mitigation
and consitute a substantial portion of planned emissions reductions
under the 2015 Paris Agreement. Yet, the efficacy of greenhouse gas
mitigation planning and reporting is dependent upon the quality of
available emission factors data, including forest carbon (C) stocks and
changes therein. Tens of thousands of relevant forest C estimates have
been published, yet are not readily accesible to the practitioners
compiling national greenhouse gas inventories. Many of these data have,
however, been compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
accounting if made available through the Emission Factor Database (EFDB)
of the International Panel on Climate Change (IPCC). Here, we develop
and document a process for semi-automated transfer of data from ForC
into the EFDB, assess the data available and transferred to date, and
provide recommendations for improving forest data collection, analysis,
and reporting to improve accounting of forest-sector greenhoouse gas
emissions and removals. We begin by reconciling terminology and mapping
ForC fields into EFDB. This process required some updates to the ForC
database structure, leading to the release of a new version of ForC
(v4.0; described here). At the time of writing, ForC contained
\textasciitilde17204 independent records that would be relevant to EFDB,
1214 of which have been submitted to date. Among the data in ForC, there
is disproportionate representation of biomass (particularly aboveground)
stocks, with far fewer records for dead organic matter and soil carbon,
and relatively few or no records for net annual increments or C fluxes
into (gains) or out of (losses) the IPCC-defined C pools. Geographic
representation is also quite uneven, with the highest densities of
relevant records in temperate forests, and with relatively scant
representation of tropical forests in Africa and Asia. ForC represents a
diversity of stand ages, although records for young stands are primarily
limited to C stocks, as opposed to net increments of fluxes. This
distribution of records is generally reflected in the subset of records
that have been submitted to EFDB to date. In the future, forest C
estimates in EFDB can be improved through targeted research to fill
critical gaps, reporting of information required by IPCC, and continued
submission of data from scientific publications to the EFDB.

@@

As of April 27, 2022, ForC (v4.0) contained 32550 independent records
(39856 total), 17204 of which were for variables relevant to EFDB (Fig.
1). These records were distributed across all forested continents and
ecozones, with particularly high concentrations in
\emph{{[}ecozone(s){]}} and low concentrations in
\emph{{[}ecozone(s){]}} (Fig. 2). The most widely represented forest
type was \emph{{[}type{]}}, followed by \emph{{[}type{]}} and
\emph{{[}type{]}} (Fig. 3).

@@

\begin{longtable}[t]{l|l|l|l|l}
\caption{\label{tab:table_variables}\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.}}\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\endfirsthead
\caption[]{\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} \textit{(continued)}}\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\endhead
\textbf{Biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass & 1095 & 847 & 93 & 48\\
\hline
delta.biomass & 0 & 0 & 0 & 0\\
\hline
NPP\_woody & 136 & 93 & 0 & 0\\
\hline
woody.mortality & 0 & 0 & 0 & 0\\
\hline
\textbf{Aboveground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_ag & 9050 & 7737 & 1251 & 693\\
\hline
biomass\_ag\_woody & 460 & 366 & 10 & 10\\
\hline
biomass\_ag\_foliage & 601 & 502 & 49 & 27\\
\hline
delta.agb & 166 & 128 & 123 & 123\\
\hline
ANPP\_woody & 299 & 242 & 0 & 0\\
\hline
woody.mortality\_ag & 112 & 62 & 30 & 17\\
\hline
\textbf{Belowground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_root & 4629 & 4180 & 123 & 55\\
\hline
biomass\_root\_fine & 931 & 594 & 18 & 18\\
\hline
biomass\_root\_coarse & 599 & 410 & 12 & 7\\
\hline
delta.biomass\_root & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_coarse & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_fine & 0 & 0 & 0 & 0\\
\hline
woody.mortality\_root & 0 & 0 & 0 & 0\\
\hline
BNPP\_root\_fine & 489 & 333 & 0 & 0\\
\hline
BNPP\_root.turnover\_fine & 91 & 56 & 0 & 0\\
\hline
BNPP\_root\_coarse & 329 & 250 & 0 & 0\\
\hline
\textbf{Dead wood} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
deadwood & 437 & 303 & 103 & 61\\
\hline
deadwood\_standing & 152 & 120 & 17 & 17\\
\hline
deadwood\_down & 424 & 368 & 51 & 27\\
\hline
delta.deadwood & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_standing & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_down & 0 & 0 & 0 & 0\\
\hline
R\_het\_deadwood & 0 & 0 & 0 & 0\\
\hline
\textbf{Litter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
O.horizon & 38 & 38 & 38 & 38\\
\hline
delta.O.horizon & 4 & 4 & 4 & 4\\
\hline
litter & 30 & 30 & 23 & 23\\
\hline
delta.litter & 0 & 0 & 0 & 0\\
\hline
NPP\_litter & 0 & 0 & 0 & 0\\
\hline
R\_het\_litter & 167 & 143 & 0 & 0\\
\hline
\textbf{Soil organic matter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
SOM / SOC & 693 & 398 & 89 & 46\\
\hline
delta.SOM / delta.SOC & 0 & 0 & 0 & 0\\
\hline
R\_het\_soil & 0 & 0 & 0 & 0\\
\hline
\textbf{TOTAL} & \textbf{20932} & \textbf{17204} & \textbf{2034} & \textbf{1214}\\
\hline
\end{longtable}
\endgroup{}

@@

As of April 27, 2022, we had reviewed or added 2034 EFDB-relevant
records, 1214 records of which were sent to EFDB, and 73 of which have
been reviewed, accepted, and posted (Figs. 2-3, Table 2).
\emph{{[}DETAILS{]}}
